Gravitational waves from colliding black holes

### 2017-04-30 Report/Draft/draft.tex
Report latest

An extremely useful and computationally cheap way to identify some of these degeneracies is through the process of matching as defined in (13). In the \textit{PyCBC} framework, the match is maximised over extrinsic parameters such as distance and source location that only affect the amplitude of a signal, and so is a useful tool for examining the similarities in the morphologies of different waveforms. A match takes values in the range $[0,1]$, with identical waveforms having a match of 1. It is also important to note that the inclination parameter has a similar nature of that of the initial phase, in that in matches of non-precessing binaries it can be maximised over as it does not affect signal morphology, however once precession is added, waveforms at different inclinations no longer match, further adding to the complexity of the parameter space. In this section we intend to get some idea of the areas in which we expect the detector to be most sensitive to precession, and those in which precessing signals are intrinsically degenerate with non-precessing waveforms. This is done by incrementally changing waveform parameters and seeing how the match between the waveforms changes as a result. 
	\includegraphics[scale=0.58]{fig9.png}
	\caption{Affect of phase on non-precessing and precessing waveforms. Precession adds an extra degree of freedom in that the morphology of precessing waveforms is different for different phases.}

@@

\begin{figure}[h]
	\includegraphics[scale=0.55]{fig10.png}
	\centering
	\caption{Matches between precessing and non-precessing waveforms for a range of inclinations and mass ratios.(make this more compact before submit)}
	\centering
\end{figure}

--Fig 11 will be spinDegen for 3 spins on 2nd BH

@@

\section{Detector response simulations}
In this section we describe the process of inserting signals into simulated detector noise, and then perform a full parameter estimation.
\subsection{Signal extraction}
Data whitening - basically LOSC stuff, how we go from raw data to a signal - move this to intro?
\subsection{Inference pipeline}
A large number of software injections were going to be necessary to generate any interesting results due to the large size and complexity of the parameter space as revealed by the match results. So it was key that they could be performed efficiently and in an organised way. This required the construction of an inference pipeline which formed the bulk of the computational work of the project. This was ultimately one bash script that would create a new folder for each run, generate the injection file and run the inference MCMC.  It also saves the injected parameters as a python dictionary, plots all posteriors overlaid with injected parameters, plots the injected waveform, and the recovered (maximum posterior, or MAP waveform) and injected waveforms overlaid on the whitened detector strain data, and return matches between the MAP and injected waveforms. This meant that all the relevant analysis and data processing needed for each inference run was fully automated. A flow chart of this process is shown in Fig. 12.
This project work was also split between a number of different LIGO clusters, as well as work on several different local PCs, and so the pipeline along with all the relevant match and precession scripts were maintained as part of a GitHub repository which meant that the codes could be developed effectively and without conflict.
\begin{figure}[h]
	\includegraphics[scale=0.55]{fig13.png}
	\centering
	\caption{Flow chart of the inference pipeline. The ellipses represent data files that were generated during the pipeline and stored in case further analysis was necessary. The boxes at the bottom represent the final output of the pipeline.}
	\centering
\end{figure}
\subsection{Inference runs}
Posteriors and discussion of inference results. Great result that injected appears to match the signal better. Weakness in the inference pipeline? Perhaps its the case that match drops here due to the fact that precessive waveforms can often only really be significantly distinguished during the peaks of the merger phase. These are obviously susceptible to random noise fluctuations. Strong conclusion here.
\subsection{Impact of Virgo}
A look at how Virgo will influence PE, especially chi p estimation.
\section{Conclusions and implications}
Summary of results on PE and estimation of chi p, and a prospect on Virgo's impact.
\section{Reflective Statement}
I now need a therapist

Still to do:

@@

\bibliography{draftbib.bib}
\end{document}

>>>>

An extremely useful and computationally cheap way to identify some of these degeneracies is through the process of matching as defined in (13). In the \textit{PyCBC} framework, the match is maximised over extrinsic parameters such as distance and source location that only affect the amplitude of a signal, and so is a useful tool for examining the similarities in the morphologies of different waveforms. A match takes values in the range $[0,1]$, with identical waveforms having a match of 1. It is also important to note that the inclination parameter has a similar nature of that of the initial phase, in that in matches of non-precessing binaries it can be maximised over as it does not affect signal morphology, however once precession is added, waveforms at different inclinations no longer match, further adding to the complexity of the parameter space.

\begin{figure}[h]
	\includegraphics[scale=0.58]{fig9.png}
	\centering
	\caption{Affect of phase on non-precessing and precessing waveforms. Precession adds an extra degree of freedom in that the morphology of precessing waveforms is different for different phases.}
	\centering
\end{figure}

In this section we intend to get some idea of the areas in which we expect the detector to be most sensitive to precession, and those in which precessing signals are intrinsically degenerate with non-precessing waveforms. This is done by finding the match between waveforms with and without in-plane spin for a range of parameters. In the case of a strong match ($>0.95$), it is unlikely that $chi_p$ can be accurately and precisely recovered for signals with those parameters due to an intrinsic similarity of the waveforms.

\begin{figure}[h]
	\includegraphics[scale=0.55]{fig10.png}
	\centering
	\caption{Matches between precessing and non-precessing waveforms for a range of inclinations and mass ratios.(make this more compact before submit)}
	\centering
	\includegraphics[scale=0.55]{fig11.png}
	\centering
	\caption{Spin difference required for the match between precessing and non-precessing waveforms to drop below 0.95}
	\centering
\end{figure}

@@

In Fig. 11, we examine the effect of the spin on the lower mass black hole on matches between precessing and non-precessing waveforms. In these figures, for a ranges of inclinations and mass ratios, we find the minimum spin difference required for the match between waveforms to drop below 0.95. This is an efficient way of quantifying the precessing degeneracy for a range of parameters. The dark regions are areas where only a small difference in spin is required for the match to drop, and the light areas show where the waveforms are degenerate. In Fig. 11 we compare results for systems with no spin, spin aligned with $\vec{L}$ and fully in-plane spin on the smaller black hole. The spin on the smaller black hole appears to only have a significant effect the matches around an inclination of 1.5, where the system is edge on, but other than that it does not drastically change the structure of the parameter space. This is convenient as it indicates that inference results for systems with $|\vec{S_2}|=0$ should be broadly applicable to those with arbitrary spins on the smaller black hole. It is also evident that the mass ratio does not appear to affect the shape of the parameter space, but more the intensity with which the match changes as a function of inclination. So we can expect the same behaviour at different mass ratios, just with the degeneracies being weaker or higher for extreme and close mass ratios respectively.

@@

Lastly we explore the impact of polarisation and phase on the parameter space, using the same technique as in Fig. 11. We select an extreme mass ratio binary here ($m1=55, m2=15$) to highlight any prominent features of the parameter space. A full range of inclinations and phases are shown for four polarisations: plus, mixed but plus dominated, mixed but cross domianted, and cross polarisations. The region of extremely high sensitivity at $\iota=pi/2$ is a result of the fact that there is no cross-polarised signal for an edge-on binary. The parameter space apperas to be highly structured, and dependent on both phase and inclination. The effect of polarisation appears to be strongest around $\pi/2$, and does not have such a large impact at other inclinations or phases.

\section{Detector response simulations}
In this section we describe the process of inserting signals into simulated detector noise, and then perform a full parameter estimation.
\subsection{Signal extraction}
Data whitening - basically LOSC stuff, how we go from raw data to a signal - move this to intro?
\subsection{Inference pipeline}
A large number of software injections were going to be necessary to generate any interesting results due to the large size and complexity of the parameter space as revealed by the match results. So it was key that they could be performed efficiently and in an organised way. This required the construction of an inference pipeline which formed the bulk of the computational work of the project. This was ultimately one bash script that would create a new folder for each run, generate the injection file and run the inference MCMC.  It also saves the injected parameters as a python dictionary, plots all posteriors overlaid with injected parameters, plots the injected waveform, and the recovered (maximum posterior, or MAP waveform) and injected waveforms overlaid on the whitened detector strain data, and return matches between the MAP and injected waveforms. This meant that all the relevant analysis and data processing needed for each inference run was fully automated. A flow chart of this process is shown in Fig. 12.
This project work was also split between a number of different LIGO clusters, as well as work on several different local PCs, and so the pipeline along with all the relevant match and precession scripts were maintained as part of a GitHub repository which meant that the codes could be developed effectively and without conflict.
\begin{figure}[h]
	\includegraphics[scale=0.55]{fig13.png}
	\centering
	\caption{Flow chart of the inference pipeline. The ellipses represent data files that were generated during the pipeline and stored in case further analysis was necessary. The boxes at the bottom represent the final output of the pipeline.}
	\centering
\end{figure}
\subsection{Inference runs}
In this section we present the results from a range of inference runs, attempt to assess the effectiveness of spin estimation and compare the inference results with the match results to evaluate how effective the match process is as a predictor of inference accuracy.
\subsection{Impact of Virgo}
A look at how Virgo will influence PE, especially chi p estimation.
\section{Conclusions and implications}
Summary of results on PE and estimation of chi p, and a prospect on Virgo's impact.
\section{Reflective Statement}
I now need a therapist

$9.98\pm0.291\times10^{-3}$
Still to do:

@@

%\bibliography{draftbib.bib}
\end{document}
